---
title: "Reducing the Risk of COVID-19 for Vulnerable ages"
author: 
  - Mohid Sharif
thanks: "Code and data are available at: https://github.com/MohidSharif/COVID19_data_paper"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "COVID-19 has taken over the world for more than two years now and it is important to study and understand in which circumstances cases rise and fall. I studied and analyzed data obtained about COVID-19 cases from various communities in Toronto specifying age, gender, and source of infection of the patient. I found that younger age groups are more likely to be infected in their household while the seniors are more likely to be infected in community settings. My findings can be used to identify circumstances in which outbreaks are less likely for certain age groups."
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
# Turn off echo for all code, disables knitting code chunks
knitr::opts_chunk$set(echo = TRUE)
# Turn off warning messaging in knitten paper file
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# Import needed librarys
library(tidyverse)
library(kableExtra)
```

# Introduction

We have been battling Covid-19 for more than two years now and we need to be strong in our battle. I wanted to read about the data from cases over Toronto and analyse why cases are still growing and how we can reduce the risk of infection and stop this virus in its tracks. Analyzing the data set we can compare variables such as age groups and gender to isolate sources of infection which are more likely to infect certain groups and find changes to those circumstances to reduce the likeliness of infection in those groups. Using this technique we can keep making adjustments to what should and should not be allowed during the pandemic to ensure cases don't rise uncontrollably.

I first analysed the data set to understand how cases had been recorded. Every case in the data set records the source of infection, age group, and gender of the patient. These are only a few of the variables but they are also the most important and they are the ones I mainly looked at in this data study. I then created two graphs to compare the amount of cases between different variables. The first graph is a comparison of cases between age groups, I found that younger ages (0-19 years) were most likely to be infected in households while seniors are most likely to be infected in community settings or healthcare institutes. The second graph compares the cases between genders, I used these graphs to analyze which gender was more likely to be infected in certain settings. I found that both males and females had almost identical data for how they were being infected, however there were twice as many females being infected in healthcare institutes.

I can use my data to isolate situations in which infections are more likely for certain groups and how we can change that to reduce infections. Firstly younger ages are more likely to get infected in their homes, in this case they are likely infected by a family member. Family members should in this case sanitize regularly while entering and exiting the house and reduce outdoor activity to necessities only. Secondly seniors are most likely to be infected in healthcare institutes because of regular visits for their health. For this, healthcare institutes should do their best to isolate the non-infected seniors from the infected reducing their risk for infection. Lastly females are more likely than men to be infected in healthcare institutes due to the fact that women dominate nursing, otherwise both males and females are just as likely to be infected in any other setting.

The remainder of this paper is: Section \@ref(data).


# Data

I obtained my data from the City of Toronto **opendatatoronto** database portal, using the 'opendatatoronto' package [@citeopendatatoronto] and the statistical programming language **R** [@citeR]. I used the **tidyverse** package for data manipulation [@citetidyverse] and **kableExtra** for table formatting [@citekableExtra]. The header includes two lines of code **"usepackage{float}"** which allows the use of float in our R markdown and the line **"floatplacement{figure}{H}"** [@citeholdposition] which keeps the tables and figures locked in the specific place where they are written in R markdown.

This data set records COVID-19 cases across Toronto and records the details of the infected. The data records the **age group, gender, neighborhood, source of infection, and if the infection was outbreak associated or sporadic**. The data classifies cases by age group rather than specific ages, age groups are grouped by 10 year age groups (E.g. "20 to 29 Years") except for the first group which is **"19 and younger"** and the last group which is **"90 and older"**. Gender is self-reported and sorted into 7 categories, from which we will look at **males** and **females**. 

Source of infection is determined by three factors:

* A public health investigator's assessment of the most likely source of infection
* Being associated with a confirmed COVID-19 outbreak
* Reported risk factors such as contact with a known case or travel

The classifications are very self explanatory categorized as household contact, close contact, community, outbreak associated (this is then further classified into 3 categories, congregate settings, healthcare institutions, and other settings), and travel.

(Table \@ref(tab:dataextract)) shows the data recorded from the neighborhood of Agincourt North. Most cases in this data set are classified as "No Information" or "Pending" under source of infection, these cases need to be discluded to see cases which give us more valuable information.



```{r dataextract, echo= FALSE}
# Import knitr
library(knitr)

# Read the data into a variable
covid19_outbreaks <- read.csv("~/starter_folder-main/inputs/data/covid19_outbreaks.csv")

# Filter the data to be specific to Agincourt North neighborhoods and to exclude source of infections that contain "No Information" or "Pending"
covid19_outbreaks |> 
  filter(Neighbourhood.Name == "Agincourt North", Source.of.Infection != "No Information", Source.of.Infection != "Pending") |>
# Pipe this to include only source of infection, age group and client gender in the table
  select(Source.of.Infection, Age.Group, Client.Gender) |>
# Include 10 listings in the table
  slice(1:10) |> 
# Display the table with a caption and appropriate column names and format properties
  kable(
    caption = "Outbreaks in Agincourt North",
    col.names = c("Source of Infection", "Age Group", "Gender"),
    digits = 1,
    booktabs = TRUE,
    linesep = "") |> 
    kable_styling(latex_options = "HOLD_position")
```

(Table \@ref(tab:dataextract)) line 1 can be read as such; a female between the ages of 30 and 39 from Agincourt North likely to be infected due to travel. The table is a good representation of which data is available to us and how we can use it to obtain which sources of infection are most common by counting all cases that fall under that source of infection. I specifically filtered this table to only include cases from Agincourt North to reduce result size and to specify to a single location in Toronto. I also filtered all sources classified as "No Information" or "Pending" since most cases in this dataset are classified as such and provides us with no information on the source of infection.

\newpage

From here I further cleaned the data for graphing. I cleaned the data by removing unnecessary or irrelevant information from columns such as "No Information" or "Pending" from source of infection. I wanted my graph to show how susceptible ages were being infected, I learnt that vulnerable ages range from the ages of 0-19 and 70+ [@citevulnerableages]. I then filtered the ages that I needed to analyze, and graphed the count data in a bar graph. 

(Figure \@ref(fig:agegraph)) shows the comparison of cases between vulnerable age groups for each source of infection.



```{r agegraph, fig.cap="Source of Infections by Age group", echo = FALSE}
# Filter the data to exclude source of infections that contain "No Information" or "Pending"
covid19_outbreaks |> 
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending") |>
# Filter further to only include ages 19 and below and 70 and above
  filter(Age.Group != "20 to 29 Years", Age.Group != "30 to 39 Years", Age.Group != "40 to 49 Years",
        Age.Group != "50 to 59 Years", Age.Group != "60 to 69 Years") |>
# Plot this graph by counting all values of source of infection for each age group
  ggplot(mapping = aes(x = Source.of.Infection, fill = Age.Group)) +
  geom_bar()+
# Format the bar graph with a theme and appropriate axis
   theme_classic()+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       y = "Cases",
       fill = "Age group")
```

The graph shows a great visual comparison of the amount of infections per age group for each source of infection. The biggest observation here is the difference in cases between the seniors and children/youth in the Household Contact category. We can see here that children/youth are much more likely to be infected in their household. Seniors are most likely to be infected in healthcare institutes likely due to their frequent visits to the hospital due to old age. The next most likely source of infection for them is community settings since seniors are often involved in community events and gatherings. Both children/youth and seniors have very low cases in the travel and congregate settings categories.

Our biggest concern from seeing this comparison should be to reduce the risk of infection in household settings for children/youth and the risk of infection in healthcare institutes. Therefore, we should try our best to sanitize before and after entering the household and to limit our outdoor activities to only necessities. For healthcare institutes we should try to keep the infected away from the other patients in the facility as much as possible. Including sanitation stations in every corridor and room, and sanitizing as often as necessary is a great way to prevent the spread of the virus.

\newpage

Now lets compare the source of infections between men and women. I used the same filters from (Figure \@ref(fig:agegraph)) for the source of infections and changed the age filters to gender filters. I then filtered out any gender classifications other than male and female since I was only interested in the two. The I proceeded to graph the amount of cases for each source of infection and graph them in a bar graph.

\@ref(fig:gendergraph)) shows the comparison of cases between males and females for each source of infection.


```{r gendergraph, fig.cap="Source of Infections by Gender", echo = FALSE}
# Filter the data to exclude source of infections that contain "No Information" or "Pending", also filter out all genders except MALE and FEMALE
covid19_outbreaks |>
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending", Client.Gender != "UNKNOWN", Client.Gender != "NON-BINARY",
          Client.Gender != "NOT LISTED, PLEASE SPECIFY", Client.Gender != "OTHER", Client.Gender != "TRANS MAN", Client.Gender != "TRANS WOMAN",
          Client.Gender != "TRANSGENDER")|>
# Plot a bar graph with counts of all cases for males and females
  ggplot(mapping = aes(x = Source.of.Infection, fill = Client.Gender)) +
# Set position of the bar to dodge so we there are multiple bars side by side
  geom_bar(position = "dodge")+
# Set theme and format the graph with appropriate axis
   theme_classic()+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       y = "Cases",
       fill = "Gender")
```


Observing the graph we can see that both males and females are very similar in the amount of cases in each source of infection. The most important information we can gather from this graph is that there are twice as many cases for women in healthcare institutes compared to men. This is most likely due to the fact that nursing is female dominated and having to take care of infected patients they have a higher chance of being infected. Other than this both males and females are most likely to be infected in community and household settings. Therefore we should encourage social distancing, regular sanitization and limit outdoor activity to only necessary activities.


\newpage
# References


