---
title: "COVID-19 Outbreaks in Toronto"
author: 
  - Mohid Sharif
thanks: "Code and data are available at: https://github.com/MohidSharif/COVID19_data_paper"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "COVID-19 has taken over the world for more than two years now. It is important to understand the measures that are being taken to explore the outbreaks throught the city. I have obtained and analyzed data from multiple COVID-19 outbreaks to compare the different variables and how they relate to eachother in different parts of the city. My findings can be used to identify circumstances in which outbreaks are less likely."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

First paragraph is motivational and broad.

Second paragraph is about what was done and what was found.

This paragraph about implications.

The remainder of this paper is: Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 



# Data

Paragraph or two introducing the dataset broadly.

Then show an extract of the dataset (Table \@ref(tab:dataextract)).

```{r dataextract, echo= FALSE}
library(tidyverse)
library(knitr)

covid19_outbreaks <- read.csv("~/starter_folder-main/inputs/data/covid19_outbreaks.csv")

covid19_outbreaks |> 
  filter(Neighbourhood.Name == "Agincourt North", Source.of.Infection != "No Information", Source.of.Infection != "Pending") |>
  select(Source.of.Infection, Age.Group, Client.Gender) |>
  slice(1:10) |> 
  kable(
    caption = "Outbreaks in Agincourt North",
    col.names = c("Source of Infection", "Age Group", "Gender"),
    digits = 1,
    booktabs = TRUE,
    linesep = "")
```


```{r dataextract2, echo= FALSE}

covid19_outbreaks |> 
  filter(Neighbourhood.Name == "York University Heights", Source.of.Infection != "No Information", Source.of.Infection != "Pending") |>
  select(Source.of.Infection, Age.Group, Client.Gender) |>
  slice(1:10) |> 
  kable(
    caption = "Outbreaks in York University Heights",
    col.names = c("Source of Infection", "Age Group", "Gender"),
    digits = 1,
    booktabs = TRUE,
    linesep = "")


```



Figure \@ref(tab: childrengraph) shows the relationship between age and source of infection.


```{r childrengraph, fig.cap="Source of Infections for vulnerable ages (ages 0-19 and 60+)", echo = FALSE}
library("gridExtra")
childrengraph <-
  covid19_outbreaks |> 
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending", Age.Group == "19 and younger") |>
  ggplot(mapping = aes(x = Source.of.Infection), stat="count") +
  geom_bar(fill = "#FF6666")+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       caption = "Youth/Children infections")

elderlygraph <-
  covid19_outbreaks |> 
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending", Age.Group == "50 to 59 Years"|| Age.Group == "60 to 69 Years"|| Age.Group == "70 to 79 Years"|| Age.Group == "80 to 89 Years"|| Age.Group == "90 and older") |>
  ggplot(mapping = aes(x = Source.of.Infection), stat="count") +
  geom_bar(fill = "#FF6666")+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       caption = "Elderly infections")

malegraph <-
  covid19_outbreaks |> 
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending", Client.Gender == "MALE") |>
  ggplot(mapping = aes(x = Source.of.Infection), stat="count") +
  geom_bar(fill = "#FF6666")+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       caption = "Male infections")

femalegraph <-
  covid19_outbreaks |> 
  filter(Source.of.Infection != "No Information", Source.of.Infection != "Pending", Client.Gender == "FEMALE") |>
  ggplot(mapping = aes(x = Source.of.Infection), stat="count") +
  geom_bar(fill = "#FF6666")+
   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  labs(x = "Source of Infection",
       caption = "Female infections")

grid.arrange(childrengraph, elderlygraph, ncol = 2)
grid.arrange(malegraph, femalegraph, ncol = 2)
```


Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


